A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2019-08-23 abstract.Rmd
revised and merged abstract

In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments. The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility. The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible. While dynamic document generation is increasingly recognized as a tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.

Reproducible Research has become the minimal standard for scientific work [@hardwicketome.DataAvailabilityReusability] in recent years. The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software and code for the researcher and her colleagues. The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them. These tools include software for literate programming, version control, dependency tracking, and containerization. If implemented, such workflow enables the researcher to transparently create, dynamically render, and publicly share reports, which she and her colleagues can trust to reproduce later, regardless of operating system or software version.

>>>>

In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses. The workflow is based on established tools and practices from software engineering. It combines the benefits of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform and long-term computational reproducibility. The workflow ensures meeting the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible. While dynamic document generation is increasingly recognized as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
